% Chapter 1

\chapter{Discussion/Conclusions} % Main chapter title

\label{Chapter4} % For referencing the chapter elsewhere, use \ref{Chapter1} 

\lhead{Chapter 4. \emph{Discussion}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Summary of Findings}
\begin{comment}
Here you will bring together the work of the dissertation by showing how the initial research plan has been addressed in such a way that conclusions may be formed from the evidence of the dissertation. 
No new material or references should be placed here. 
The conclusions should make a statement on the extent to which each of the aims and objectives has been met. 
You should bring back your research questions and state clearly your understanding of those questions.
Be careful not to make claims that are not substantiated from the evidence you have presented in earlier chapters.
If you are undertaking a company project based around a business issue do not confuse recommendations for the company with conclusions. 

Provide inferences and implications that the results of the study provide you and
the reader or others who may have interest in the results. This is a time to
expound on your results and offer insight into what your study does or does not
contribute to the body of information on your topic.

Conclusions Drawn by Results
Identify specific conclusions resulting from you study. Offer specific insight to
what your findings reveal. This section should synthesize your findings with the
current knowledge in your area of study.



\section{Recommendations for Further Research}
Provide recommendations to further research on this topic or how parts of your
study could be improved upon. If you found as a result of your study that
another topic should be looked at in order to offer more insight into this topic,
then suggest that at this time. It is important that this part of your conclusion
chapter incorporate the implications of your findings in terms of other research in
your area of study.

If you want to include a list of recommendations then do so in a separate short chapter. 
The conclusions address the wider understanding of the issue you have been studying.
You should include a short sub section on any suggestions for further research for colleagues who might wish to undertake research in this area in the future.
There should also be a short statement of the limitations of the research. 
Often as a single case study or limited range of companies you can not really claim
that your research holds for all companies. 
However, by adopting a rigorous approach to your literature review and methods which have validity and can be repeated you can make a reasonable but limited claim that your conclusions
should be taken seriously.
\end{comment}



